### 2020-04-21 Thesis-Chapter2-Background.tex
Finish EA types brief overview

% There are several types of approaches:
% TODO: my selections here are also rather sparse. Create another appendix? (I love work). Possibly create table of
% properties below, similar to denning_taxonomy_1996
% - Allow strong encryption (anyone with the users' private keys can see it)
% - Disallow encryption (anyone can see it)
% - Make encryption sufficiently weak, e.g. limit key length (anyone with enough CPU can see it)
% - Require use of backdoored algorithms, security via obscurity (anyone with the secret knowledge can see it)
% - Allow any encryption, just be able to provide keys (anyone with the extra keys can see it)
%   - Store extra in gov't DBs, in corporate DBs, in third-party DBs, on blockchain, in secure HW on device
%   - Split the key and store it multiple places
%   - See denning_taxonomy_1996
% - Use some alternative path
%   - Coerced passwords, finding unencrypted copies, seizing devices while unlocked, lawful hacking

@@

@@

\cite{denning_taxonomy_1996}
\ac{clipperchip} \cite{blaze_protocol_1994}
% The technical approach used by the \ac{clipperchip} secret protocol hardware enabled split-key government escrow.

@@

\cite{goos_oblivious_1996}
\cite{phan_key_2017}

@@

\cite{savage_lawful_2018}
\cite{ozzie_2018}

@@

\propsstart
    \prop{Translucent Cryptography}{

\cite{bellare_translucent_1996}

@@

\cite{nguyen_lawful_2017}
\cite{soesanto_2018}
\cite{kerr_encryption_2017}

@@

\cite{bittenbender_2019}
\cite{kerr_encryption_2017}

@@

% access device while unlocked; find copies elsewhere
\cite{kerr_encryption_2017}

>>>>

% There are several types of approaches:
% TODO: my selections here are also rather sparse. Create another appendix? (I love work). Possibly create table of
% properties below, similar to denning_taxonomy_1996. savage_lawful_2018 has some good citations to mine.
% - Allow strong encryption (anyone with the users' private keys can see it)
% - Disallow encryption (anyone can see it)
% - Make encryption sufficiently weak, e.g. limit key length (anyone with enough CPU can see it)
% - Require use of backdoored algorithms, security via obscurity (anyone with the secret knowledge can see it)
% - Allow any encryption, just be able to provide keys (anyone with the extra keys can see it)
%   - Store extra in gov't DBs, in corporate DBs, in third-party DBs, on blockchain, in secure HW on device
%   - Split the key and store it multiple places
%   - See denning_taxonomy_1996
% - Use some alternative path
%   - Coerced passwords, finding unencrypted copies, seizing devices while unlocked, lawful hacking

@@

% TODO: need an aside on shamir m/n secret sharing? Probably in order to understand split key approaches at all

@@

Any scheme that relies on the fidelity and security of information held by an entity besides the data owner (in the case
of \ac{DAR}) or sender or receiver (in the case of \ac{DIM}) is a ``trusted-party'' key escrow approach. Though there is
a wide variety of implementation options within this category \cite{denning_taxonomy_1996}--secret or open protocol,
hardware-enabled or software-only, single key or split key, government or third-party escrow--the foundation of the
effectiveness of these approaches lies in the security of the trusted party. The \ac{clipperchip} is one example, being
a secret protocol hardware-enabled split key government escrow system \cite{denning_taxonomy_1996}.

@@

Some schemes compensate for the risk of concentrated sensitive information by massively distributing the secret
information. In this arrangement, trust is laid in the distrubuted system, not in any single party. Such systems offer
high data availability while making it difficult for access requests to be secret or to subvert access policies. This
approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has come into its own through the application of
blockchain \cite{phan_key_2017}. These approaches are not as developed as trusted-party \ac{key-escrow}.

@@

A third type of \ac{key-escrow} is device-based. In device-based \ac{key-escrow}, special key information is held not by
trusted parties or distributed systems, but on the device itself in specialized hardware. Advances in secure hardware
``enclaves'' since \ac{the-cw1} have enabled secure local key storage. Approaches relying on hardware primarily target
\ac{DAR}, and include time elements and split keys with trusted parties as well \cite{savage_lawful_2018}
\cite{ozzie_2018}.

@@

\propsstart
    \prop{Translucent cryptography}{

Research in \ac{the-cw1} included the introduction of ``translucent'' cryptography, in which an observer could surveil
some but not all communications using statistical cryptographic methods \cite{bellare_translucent_1996}. Under this
scheme, no keys are escrowed, but law enforcement could recover a predetermined fraction--say, 40\%--of plaintext data.
That fraction would be designed to balance privacy, digital security, and public security. This research, though
decidedly more experimental than key escrow proposals, was explicitly pursued to demonstrate that alternatives to key
escrow do exist \cite{bellare_translucent_1996}.

@@

As introduced in \mysec{sec-reg-environment}, one workaround to direct \ac{EA} is \ac{lawful-hacking}. Instead of having
access to plaintext data via an alternative means of decryption, law enforcement is allowed to perform otherwise-illegal
hacking activity in order to find the key, compromise a device, or intercept plaintext versions of the data
\cite{kerr_encryption_2017}. While some see this as a viable middle ground \cite{nguyen_lawful_2017}, encouraging law
enforcement to exploit vulnerabilities creates mis-aligned incentives. If law enforcement opposes strong security in
general, it could foster the exploit market while costing ordinary users money and security \cite{soesanto_2018}.

@@

In some cases, a suspect in custody may have the ability to access the desired encrypted information via key, \ac{PIN},
or password. In these cases, authorities sometimes compel the suspect to disclose the password. In the U.S., this
approach draws debates about the fifth amendment right against self-incrimination \cite{bittenbender_2019}. Of course,
this approach cannot address \ac{DIM} or situations in which the knowledgeable party is not in custody.

% these look good: sacharoff_unlocking_2018, fredona_fifth_2016
% this probably has something to say: kerr_encryption_2017

@@

Simpler methods can be used than computer hacking or legal coercion: encrypted data may exist in plaintext elsewhere, or
authorities may seize equipment while it is unlocked \cite{kerr_encryption_2017}. This approach relies on skills law
enforcement is traditionally good at--investigation and sting operations--and is used successfully. In 2017, an
international law enforcement operation used this technique to take down a large illicit online drug market
\cite{greenberg_2018}. This approach also does not address \ac{DIM}, and requires well-executed operations.
